---
title: "Project 1_Junxiang Yang,Viraj Upadhyay, Abheeshta Ongole, Ravi Prasad"
output: word_document
---
```{r}
#Set up the environment and load the data into R.
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(ggplotlyExtra)
library(plyr)
library(readxl)
library(repr)
library(treemapify)
library(tidyverse)
library(stringr)
library(lubridate)
library(corrplot)
library(plotly)
library(gapminder)

setwd("E:/Human Factors Engineering/Fall 21' Semester/Data Computation and Visualization/Project/fma_metadata/")
signature.data <- read.csv("echonest.csv")
df<-read.csv("tracks.csv")
```





```{r}
#Change the column name and remove the Na values.

df<- df[-c(2),]
 colnames(df)<-c("track_id","comments", "date_created", "date_released", "engineer", "favorites", "id", "information", "album_listens", "producer", "tags", "albul_title", "tracks", "type", "active_year_begin", "active_year_end", "associated_labels", "bio", "comments2", "date_created", "favorites", "id", "latitude", "location", "longitude", "members", "name", "related_projects", "tags", "website", "wikipedia_page", "split", "subset", "bit_rate", "comments3", "composer", "date_created", "date_recorded", "duration", "favorites", "genre_top", "genres", "genres_all", "information", "interest", "language_code", "license", "songs_listens", "lyricist", "number", "publisher", "tags", "songs_title")
df<- df[-c(1),]


```

###Display the songs released each year from data set
```{r}
df2 <- df[!is.na(df$date_released),]

#New table
Tracks_date<-select(df2,track_id,date_released)
typeof(Tracks_date$date_released)
num<-as.numeric(Tracks_date$date_released)


Tracks_date$date_released<-as_datetime(Tracks_date$date_released, tz="EST")


#NUMBER OF SONGS RELEASED EVERY YEAR

track_year2<-Tracks_date%>%separate('date_released',c('year','month','date'),sep='-')

track_year3<-count(track_year2,vars = 'year')
track_year3<-track_year3%>%
  arrange(desc(freq))
track_year3<-track_year3[-1,]
track_year4<-head(track_year3,25)

#BAR PLOT TO REPRESENT THE ABOVE DATA
ggplot(data=track_year4,aes(x=year,y=freq), na.rm=TRUE) +
  geom_bar(stat = "identity",position="dodge",fill="burlywood", color="darkred",na.rm=TRUE)+ggtitle("Number of songs released per year")+
   ylab("Number of songs") +
   xlab("Year")+  geom_hline(yintercept = mean(track_year4$freq), color="blue")+
   theme(plot.title = element_text(face="bold"),axis.text.x = element_text(angle = 45)) 

```

###Display zoom up view of the songs released from 2004-2013
```{r}

df3<-subset(Tracks_date, year(Tracks_date$date_released)=="2009")
p1<-length(df3$track_id)
df5<-subset(Tracks_date, year(Tracks_date$date_released)=="2011")
p2<-length(df5$track_id)
df4<-subset(Tracks_date, year(Tracks_date$date_released)=="2008")
p3<-length(df4$track_id)
df6<-subset(Tracks_date, year(Tracks_date$date_released)=="2010")
p4<-length(df6$track_id)
df7<-subset(Tracks_date, year(Tracks_date$date_released)=="2006")
p5<-length(df7$track_id)
df8<-subset(Tracks_date, year(Tracks_date$date_released)=="2004")
p6<-length(df8$track_id)
df9<-subset(Tracks_date, year(Tracks_date$date_released)=="2012")
p7<-length(df9$track_id)
df10<-subset(Tracks_date, year(Tracks_date$date_released)=="2013")
p8<-length(df10$track_id)


years_df<-c("2004","2006","2008","2009","2010","2011","2012","2013")
number_of_songs<-c("1347","1693","4072","6145","6776","7393","7633","7913")
years_songs<-data.frame(years_df,number_of_songs)

#POINT PLOT TO REPRESENT THE DATA
ggplot(data=years_songs, aes(x=years_df,y=number_of_songs), na.rm=TRUE) +
   geom_point(stat = "identity", color="darkred",na.rm=TRUE)+ggtitle("Number of songs released per year")+
   ylab("Number of songs") +
   xlab("Year")+stat_summary(fun=mean, geom="point", shape=18,size=3, color="red")
   theme(plot.title = element_text(face="bold")) 
   
```


###Display Top 10 places with most songs from the data set
```{r}
df$songs_listens<-as.numeric(df$songs_listens)

#To find the top 10 places with the most songs

df_ravi_top_origins <- select(df[!is.na(df$location),],location,track_id)

df_ravi_top_origins$location <- gsub('Portland OR', 'Portland, OR', df_ravi_top_origins$location)

df_ravi_top_origins$location <- gsub('Brooklyn, NY', 'New York, NY', df_ravi_top_origins$location)

df_ravi_top_origins$location <- gsub('Perm, Russia', 'Russia', df_ravi_top_origins$location)

df_ravi_top_origins<-
  group_by(df_ravi_top_origins,location)
df_ravi_top_origins<-count(df_ravi_top_origins,vars = 'location')
df_ravi_top_origins<-df_ravi_top_origins[-1,]

df_ravi_top_origins<- df_ravi_top_origins%>%
  arrange(desc(freq))%>%
  head(10)



ggplot(data=df_ravi_top_origins, aes(x=location,y=freq), na.rm=TRUE) +
  geom_bar(stat = "identity",position="dodge",na.rm=TRUE, fill = 'pink')+ggtitle("Top 10 places with the highest song releases")+
  ylab("Number of songs") +
  xlab("Origin") +
  theme(plot.title = element_text(face="bold"))+theme(plot.title = element_text(face="bold"),axis.text.x = element_text(angle = 270))
```
### Displaying the number of listens of Top 10 songs from the data set
```{r}
#Getting the top 10 songs from the list
df$songs_listens<- as.integer(df$songs_listens)
top10<-df[order(df$songs_listens),]
top10 <- tail(top10, n=10)

top10_ravi<-top10%>%
  select(track_id,songs_title,songs_listens)

#Plot the data
ggplot(data=top10_ravi, aes(track_id,y=songs_listens), na.rm=TRUE) +
  geom_bar(stat = "identity",position="dodge",na.rm=TRUE, fill = 'purple')+ggtitle("Top 10 songs")+
  ylab("Number of Listens") +
  xlab("Song ID") +
  theme(plot.title = element_text(face="bold"))



```


###Comparing the genres distribution in two different region
```{r}
#Case 1
#Japan Vs. UK

#Finding all the songs that are created in Japan and UK
JP<- df[df$location=='Japan',]
UK<- df[df$location=='UK',]

#Find the genres and count the frequency of each genres
JP_g<-count(JP,vars='genre_top')
UK_g<-count(UK,vars = 'genre_top')


#Clean the blank value in the data
UK_g<-UK_g[-1,]
JP_g<-JP_g[-1,]

#Create the treeplot
P_JP<-ggplot(JP_g, aes(area = freq, fill = genre_top,label= genre_top)) +geom_treemap() + geom_treemap_text(fontface = "italic", colour = "white", place = "centre",
                    grow = TRUE)+scale_fill_discrete(name='Top Genres in Japan')
P_UK<-ggplot(UK_g, aes(area = freq, fill = genre_top,label=genre_top)) +geom_treemap()+ geom_treemap_text(fontface = "italic", colour = "white", place = "centre",
                    grow = TRUE)+scale_fill_discrete(name='Top Genres in UK')

grid.arrange(P_JP,P_UK,ncol=1,top='Popular genres of music in UK and Japan')



```






###Pie chart showing the 5 most polular genres from the Top 10 songs
```{r}
#Creating a data frame showing genres of tracks and track id
track_g<-data.frame(col1=top10$track_id,col2=top10$genres)

#Manually count the frequency of each genres id shows up in the list and create a data frame

track_g1<-data.frame(col1=c('10','12','15','17','18','21','66','184','240','267','297','468','495','1235'),col2=c('2','2','5','2','2','1','2','1','1','1','1','1','1','3'))

#Sort out the top 5 genres from top 10 songs genres
track_g1<-track_g1%>%
  arrange(desc(col2))%>%
  head(5)

#Add the corresponding genre names to the genre id
track_g2<-cbind(track_g1,data.frame(c('Electronic','Instrumental','Pop','Rock','Folk')))
colnames(track_g2)<-c('genres id','count','genres_names')
track_g2$count<-as.numeric(track_g2$count)

#Calculate the percentage of each genre
all_genre<-sum(track_g2$count)
track_g2<-track_g2%>%
  mutate(perc= track_g2$count/all_genre*100)
track_g2$perc<-round(track_g2$perc,1)
track_g2$perc<-paste(track_g2$perc,'%')


#Make the pie plot 
ggplot(track_g2, aes(x = "", y = count, fill = genres_names)) +
  geom_col() +
  geom_text(aes(label = perc),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y")+ggtitle('The Top 5 genres from Top 10 songs')+scale_fill_discrete(name='Genres Name')

```


```{r}
tfy <- signature.data[1:500,27:250]
ttfy <- (t(tfy))
t <- data.frame(SrNo = 1:224, ttfy)
p1 <- ggplot(t, aes(x=SrNo, y=X2)) + geom_line() + ggtitle("Plot of Track ID vs Temporal Features")
p2 <- ggplot(t, aes(x=SrNo, y=X3)) + geom_line()+ggtitle("Plot of Track ID vs Temporal Features")
p3 <- ggplot(t, aes(x=SrNo, y=X4)) + geom_line()+ggtitle("Plot of Track ID vs Temporal Features")
p4 <- ggplot(t, aes(x=SrNo, y=X5)) + geom_line()+ggtitle("Plot of Track ID vs Temporal Features")
grid.arrange(p1,p2,p3,p4)
p5 = ggplot() +
geom_line(data = t, aes(x = SrNo, y = X2), color = "blue") +
geom_line(data = t, aes(x = SrNo, y = X4), color = "red") +
xlab("Number of Features") +
ylab("Temporal Features") +
ggtitle("Signature of Song 2 v 4") +
theme_dark()

p6 = ggplot() +
geom_line(data = t, aes(x = SrNo, y = X2), color = "blue") +
geom_line(data = t, aes(x = SrNo, y = X5), color = "red") +
xlab("Number of Features") +
ylab("Temporal Features") +
ggtitle("Signature of Song 2 v 5")


#ggplotly(p5)
#ggplotly(p6)
grid.arrange(p5,p6)
```

###Comparsion of audio features of songs from the data set and sactter plot of Danceability and Valence of the song
```{r}

trackinfo <- select(df, track_id,albul_title)
colnames(trackinfo) <- c('trackid','trackname')
trackinfo <- trackinfo %>%
  mutate(trackid = as.numeric(trackid))
Valence <- signature.data[1:2000,9]
Danceability <- signature.data[1:2000,3]
Acousticness <- signature.data[1:2000,2]
Energy <- signature.data[1:2000,4]
Instrumentalness <- signature.data[1:2000,5]
Liveness <- signature.data[1:2000,6]
Speechiness <- signature.data[1:2000,7]

Tempo <- signature.data[1:2000,8]
Genres <- df[1:2000,41]

trackid <- signature.data[1:2000,1]
trackname <- signature.data[1:2000,15]
valdan <- data.frame(Valence, Danceability, Acousticness, Energy, Instrumentalness, Liveness, Speechiness, Tempo)
valdan1 <- data.frame(Valence,Danceability,trackid,trackname)
bpm <- data.frame(Tempo,trackid,Genres)
features <- full_join (valdan1, trackinfo)

rel <- ggplot(valdan1, aes(label3=trackname)) + geom_point(aes(x=Valence, y=Danceability), color="Blue")
ggplotly(rel)

#ggplotly(rel)#Not working showing error could not find function "ggplotly" 
```
# correlation plot of different audio features
```{r}
colnames(valdan) <- c("Valence", "Danceability", "Acousticness", "Energy", "Instrumentalness", "Liveness", "Speechiness", "Tempo")
corres1<-na.omit(valdan)
M<-cor(corres1)
corrplot(M, method = 'number')

```

###Beats per minutes Vs. different tracks with Rock Genres from the data set
```{r}
bpm1<-bpm[bpm$Genres=="Rock",]

ggplot(bpm1, aes(x = trackid, y = Tempo ,fill = Tempo)) + 
    geom_bar(stat="identity") +    
    
    # To use a polar plot and not a basic barplot
    coord_polar(theta = "y") +    
    
    #Remove useless labels of axis
    xlab("") + ylab("") +
    
    ggtitle("Plot of Tracks with Rock Genres vs Tempo") +
    #Remove useless legend, y axis ticks and y axis text
    theme(legend.position = "none" , axis.text.y = element_blank() , axis.ticks = element_blank())
```

###Creating a density graph of energy audio feature being the density and x axis being the top 10 songs
```{r}

df$songs_listens<- as.integer(df$songs_listens)
top13<-df[order(df$songs_listens),]
top13 <- tail(top13, n=100)

top13_ravi<-top13%>%
  select(track_id,songs_title,songs_listens)
trackinfo <- signature.data[,1:9]
trackinfo <- trackinfo %>%
mutate(track_id = as.character(track_id))

#v1<-ggplot(tracksubset, aes(x=track_id, y=echonestenergy, fill=Albumname)) + 
#  geom_boxplot()
#v1
violinplot <- full_join(top13_ravi, trackinfo)
violinplot <- violinplot %>%
  drop_na()

d1 <- ggplot(violinplot, aes(x=echonestliveness)) + 
  geom_density()+ggtitle('Density graph of Liveness of the song')+geom_vline(aes(xintercept=mean(echonestliveness)),
            color="blue", linetype="dashed", size=1)

d2 <- ggplot(violinplot, aes(x=echonestacousticness)) + 
  geom_density()+ggtitle('Density graph of Acousticness of the song')+geom_vline(aes(xintercept=mean(echonestacousticness)),
            color="blue", linetype="dashed", size=1)

d3 <- ggplot(violinplot, aes(x=echonestspeechiness)) + 
  geom_density()+ggtitle('Density graph of Speechiness of the song')+geom_vline(aes(xintercept=mean(echonestspeechiness)),
            color="blue", linetype="dashed", size=1)

grid.arrange(d1,d2,d3,ncol=1)

```

